---
title: "Australia's Greenhouse Gas Emission Dashboard"
output: 
  flexdashboard::flex_dashboard:
    theme: sandstone
    #storyboard: true
    #orientation: rows
    orientation: columns
    #vertical_layout: scroll
    #social: [ "twitter", "facebook", "menu"]
    #source_code: embed
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(flexdashboard)
library(knitr)
library(ggplot2)
library(plotly)
library(dplyr)
library(openintro)
library(highcharter)
library(ggvis)
library(leaflet)
library(rgeos)
library(maptools)
library(mapproj)
library(rgdal)
library(ggmap)
library(broom)
library(tidyr)
library(scales)
library(RColorBrewer)
library(htmltools)
library(plyr)
```


```{r}
setwd("C:/Users/saurabh/Desktop/Data Vis/Data Emission")

```
Dashboard
=====================================  
Increasing carbon footprint is the biggest concern of this century and Australia ranks amongst the major carbon emitters, with an average per capita footprint of around 17 tonnes. This dashboard helps the viewer to understand, where does Australia stand in terms of its carbon emission and what policy makers can do to reduce country's carbon footprint. These plots are interactive, so the emission values can be viewed by hovering the pointer over plots, including the first plot, where major carbon emitters can be compared against Australia with their emission quantities.


Row {data-height=600, .bgred}
-------------------------------------
### Australia has one of the highest per capita emissions of carbon dioxide in the world, with its 0.3% of the world's population releasing 1.07% of the world's greenhouse gases.

```{r message=FALSE, warning=FALSE}

# Read data into annual_emission object
annual_emission <- read.csv('annual-co-emissions-by-region.csv') 
annual_emission$emissions_in_tonnes<-round(annual_emission$emissions_in_tonnes/1000000, digits = 2)


# Read data into per_capita_emission object
per_capita_emission <- read.csv('co-emissions-per-capita.csv') 
per_capita_emission$Per_capita_emissions_in_tonnes <- round(per_capita_emission$Per_capita_emissions_in_tonnes, digits = 2)
# Merging two dataframes by country name
emission<-merge(annual_emission,per_capita_emission, sort = TRUE, by = c("Entity","Year", "Code"))

# Read data into population object
population <- read.csv('datasets_507962_1091873_population_by_country_2020.csv')
# sort in decresing order for population
population_sorted <- population[order(-population$Population..2020.),]
# put the country's name as Entity
population_sorted$Entity <- population_sorted$Country..or.dependency.
# filter top 100 populated countries
population_sorted_top100 <- slice_max(population_sorted, population_sorted$Population..2020., n=100)
# Merge to emission dataset, so we filtered the emission data only for 100 most populated country
emission_pop<-merge(emission,population_sorted_top100, by = 'Entity')
# sort the above df yearwise
emission_pop_sorted <- emission_pop[order(emission_pop$Entity,emission_pop$Year),]

# Filter only year 2017 data, to get the most recent per capita emission countries
emission_pop_sorted_2017<- emission_pop_sorted %>% filter (Year == "2017")
# Filter the top 10 highest per capita emission producing country
emission_pop_sorted_2017_top10<-slice_max(emission_pop_sorted_2017, emission_pop_sorted_2017$Per_capita_emissions_in_tonnes, n=10 )
# Join back to 1 step previous dataset to get emission rate from previous years
emission_pop_sorted_top10<- subset(emission_pop_sorted, Entity %in% emission_pop_sorted_2017_top10$Entity)

#emission_pop_sorted_top10$emission_in_tonnes<-emission_pop_sorted_top10$emission_in_tonnes/1000000
emission_pop_sorted_top10_30yrs<- emission_pop_sorted_top10 %>% filter (Year >= 1990 & Year<2017)


emission_pop_sorted_top10_30yrs$Entity <- revalue(emission_pop_sorted_top10_30yrs$Entity, c("United Arab Emirates"="UAE"))

# Line Plot for top 10 per capita emission countries
f1 <- list(
  family = "Old Standard TT, serif",
  size = 10
  #color = "black"
  )

m <- list(
  l = 30
  #r = 50,
  #b = 100,
  #t = 100,
  #pad = 4
)

xaxis <- list(title = "",
             showgrid = FALSE,
             zeroline = FALSE,
             showline = FALSE,
             dtick = 5
             #showticklabels = FALSE
             )
yaxis <- list(title = "Metric Tonnes",
             showgrid = FALSE,
             zeroline = FALSE,
             showline = FALSE,
             titlefont = f1,
             opacity = 0.1
             #showticklabels = FALSE
             )
legend <- list(itemsizing='constant',
               orientation = 'v',
               font = list(family = "sans-serif", 
                          size = 8, color = "#000")
               
             )


emission_pop_sorted_top10_30yrs %>%
highlight_key(~Entity) %>%
  plot_ly(x = ~Year, y = ~Per_capita_emissions_in_tonnes,
              color = ~Entity, type = "area", mode = "lines", 
              colors = "Paired", reversescale=TRUE, hoverinfo = 'text',
              text = ~paste("Country: ", Entity, "<br>Annual Emission: ", emissions_in_tonnes, 'Mil.Tn<br>Per Capita Emission: ',Per_capita_emissions_in_tonnes,
                            "Mt Tn<br>Year: ", Year)) %>% 
  layout(
    #xaxis = list(zeroline = FALSE, title = "Year"),
    #yaxis = list(zeroline = FALSE, title = "Emission volume"),
    showlegend = TRUE,
    xaxis = xaxis,
    yaxis = yaxis,
    margin = m,
    legend = legend
    
  )  %>%
  highlight(on = "plotly_click", off = "plotly_doubleclick",persistent = getOption("persistent", FALSE),
  dynamic = FALSE,color = NULL,selectize = FALSE, defaultValues = "Australia", opacityDim = .2, selected = attrs_selected(showlegend = FALSE))%>%
  config(displayModeBar = FALSE)




```


-----------------------------------------------------------------
### Australia's emissions have been driven by electricity sector, followed by agriculture and transportation.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# Read data into per_capita_emission object
emissions_by_sector <- read.csv('ghg-emissions-by-sector.csv') 
# Filter dataframe for Australia 
emissions_by_sector_Aus <- emissions_by_sector  %>% filter(Entity == "Australia")
emissions_by_sector_Aus$Electricity<- emissions_by_sector_Aus$Electricity...Heat
emissions_by_sector_Aus$Manufacturing<- emissions_by_sector_Aus$Manufacturing.energy

#Collapse dataframe
emissions_by_sector_Aus_gathered<- emissions_by_sector_Aus %>%
  select(Year, Agriculture, Manufacturing,
         Transport, Electricity, Fugitive.emissions) %>%
  gather(key = "Sector", value = "Emission_in_tonnes", -Year)

emissions_by_sector_Aus_gathered$Emission_in_Mil._Tn<-emissions_by_sector_Aus_gathered$Emission_in_tonnes/1000000
## Area Plot 
# Give a specific order to all sectors:
emissions_by_sector_Aus_gathered$Sector <- factor(emissions_by_sector_Aus_gathered$Sector , levels=c("Electricity","Agriculture", "Transport", "Manufacturing",
                                                           "Fugitive.emissions"))
options(scipen = 999)
ggplotly(ggplot(emissions_by_sector_Aus_gathered, aes(fill=Sector, y=Emission_in_Mil._Tn, x=Year)) +
           geom_area(position="stack", stat="identity") +
          # annotate("text", x=2008, y=400, label= "Electricity" , 
            #         colour = "black", size = 3) +
           #annotate("text", x=2008, y=250, label= "Agriculture" , 
            #          colour = "black", size = 3) +
          # annotate("text", x=2008, y=100, label= "Transport" , 
            #         colour = "black", size = 3) +
          # annotate("text", x=2008, y=50, label= "Manufacturing" , 
            #        colour = "black", size = 3) +
           #annotate("text", x=2008, y=10, label= "Fugitive Emission" , 
            #        colour = "black", size = 3) +
           
           labs(fill = "Sector")+
           xlab("Year")+
           ylab("Million Tonnes")+
           #labs(title= "Breakdown of Australia's carbon footprint")+
           scale_fill_brewer(palette="Set2", direction = 1 ) +
           scale_x_continuous(breaks = seq(1990, 2016, by = 2))+
           scale_y_continuous(breaks = seq(0, 600, len = 7))+
           theme( 
                  legend.title =element_blank(),
                  legend.text = element_text(size = 6),
                  legend.position = "top",
                  
                  axis.title.x = element_blank(),
                  axis.title.y= element_text(family = "Old Standard TT, serif", size = 7),
                  axis.ticks.x = element_blank(),
                  axis.ticks.y = element_blank(),
                  #panel.grid.major = element_blank(), 
                  #panel.grid.minor = element_blank(),
                  plot.background = element_rect(fill = 'white'),
                  panel.background = element_rect(fill = 'white')
           )
)%>%
  config(displayModeBar = FALSE)
```


*** 

Some commentary about Frame 1.

Row {data-height=450, .bgred}
-------------------------------------
### Australia's maximum energy is generated from fossil fuels, coal accounts for majority of its total electricity generation, well above OECD average.

```{r}
# Read data into per_capita_emission object
fuel_mix <- read.csv('fuel_mix_energy.csv') 
fuel_mix_df <- as.data.frame(t(fuel_mix), stringsAsFactors = FALSE)

# Remove first row
fuel_mix_df <- fuel_mix_df[-1, ] 

# Set column names
fuel_mix_df <- setNames(cbind(rownames(fuel_mix_df), fuel_mix_df, row.names = NULL), 
                        c("Fuel.mix", "OECD_avg.", "Australia"))

fuel_mix_df$OECD_avg. <- as.numeric(fuel_mix_df$OECD_avg.)
fuel_mix_df$Australia <- as.numeric(fuel_mix_df$Australia)


# gather columns in single column
fuel_mix_gathered <- fuel_mix_df %>% 
  select(Fuel.mix,OECD_avg.,Australia) %>% 
  gather(key = "Key", value = "Percentage", -Fuel.mix)

fuel_mix_df$Fuel.mix <- factor(fuel_mix_df$Fuel.mix , 
                                    levels=c("Coal.and.Peat", "Petroleum", "Gas"))

fuel_mix_df$Fuel.mix <- revalue(fuel_mix_df$Fuel.mix, c("Coal.and.Peat"="Coal and Peat"))

f1 <- list(
  family = "Old Standard TT, serif",
  size = 10
  #color = "black"
  )

xaxis <- list(title = "",
              showgrid = FALSE,
              zeroline = FALSE,
              showline = FALSE
              
              #showticklabels = FALSE
)
yaxis <- list(title = "Percentage",
              showgrid = FALSE,
              titlefont = f1,
              zeroline = FALSE,
              showline = FALSE
)

m <- list(
  l = 30
  #r = 50,
  #b = 100,
  #t = 100,
  #pad = 4
)


  fig <- plot_ly(fuel_mix_df, x = ~Fuel.mix, y = ~Australia, type = 'bar', name = 'A',  colors = "#31a354",  opacity=0.80,                      hoverinfo = 'text',
                 text = ~paste("Australia: ", Australia, "%<br>"), textposition = 'auto',
                 marker = list(color = '#31a354',
                               line = list(color = '#31a354', width = 1.5)))
  fig <- fig %>% add_trace(y = ~OECD_avg., name = 'O',hoverinfo = 'text', text = ~paste("OECD Avg.: ", OECD_avg., "%<br>"), textposition = 'auto', 
                           colors = "#D95F02",
                           marker = list(color = '#D95F02',
                                         line = list(color = '#D95F02', width = 1.5)))
 
  fig <- fig %>% layout(yaxis = yaxis, xaxis = xaxis, barmode = 'group', showlegend = FALSE, margin = m)%>%
  config(displayModeBar = FALSE)
  fig



  
```


### Animal agriculture is the leading cause of global warming, it emits about 50% of all greenhouse gases, over 20 years, more than all other sources.

```{r}
# Read data into per_capita_emission object
Agriculture_emission_all <- read.csv('Agriculture emission.csv') 

 
Agriculture_emission_sliced<-Agriculture_emission_all[1:8,1:5]
Agriculture_emission_final <- slice_max(Agriculture_emission_sliced, Total.CO2.eMt., n=5)
#Agriculture_emission_sliced_sorted <- Agriculture_emission_sliced[order(-Agriculture_emission_sliced$Total.CO2.eMt.),]
Agriculture_emission <- Agriculture_emission_final[ , 1:4]


Agriculture_emission$Greenhouse.gas.source <- revalue(Agriculture_emission$Greenhouse.gas.source, c("Enteric fermentation"="Livestock", "Prescribed burning of savanna"="Burning Savana", "Manure management" = "Manure", "Agricultural soils"="Soils", "Urea application" = "Urea"))
Agriculture_emission$CO2<-Agriculture_emission$Carbon.dioxide.CO2.eMt.
Agriculture_emission$Methane<-Agriculture_emission$Methane.CO2.eMt.
Agriculture_emission$Nitrous.oxide<-Agriculture_emission$Nitrous.oxide.CO2.eMt.
#Collapse dataframe
Agriculture_emission_gathered<- Agriculture_emission %>%
  select(Greenhouse.gas.source, CO2, Methane, Nitrous.oxide) %>%
  gather(key = "Gas", value = "Emission_CO2_equivalent_Mt", -Greenhouse.gas.source)

Agriculture_emission_gathered$Greenhouse.gas.source<- reorder(Agriculture_emission_gathered$Greenhouse.gas.source, -Agriculture_emission_gathered$Emission_CO2_equivalent_Mt)

ggplotly(ggplot(Agriculture_emission_gathered, aes(fill=Gas, y=Emission_CO2_equivalent_Mt, x=Greenhouse.gas.source)) +
  geom_bar(stat="identity") +
  #stat_summary(aes(label = stat(y)), fun= 'sum', geom = 'text', col = 'black', vjust = 1.5) +
  #stat_summary(fun.y = sum, aes(label = ..y.., group = Gas), geom = "text")
  labs(fill = "Gas")+
 # xlab("Greenhouse gas source")+
  ylab("CO2.e Mt")+
 # labs(title= "Breakdown of Emission from Agriculture Sector")+
  #geom_text(aes(label = c("66.3%","15.5%","10.8%","3.9%","1.5%","","","","","","","","","","" )),
    #size = 3, hjust = 0.5, vjust = 0, position = "stack")+
  #geom_text(aes(label=Greenhouse.gas.source),vjust=0, size=10)+
  scale_y_continuous(breaks = seq(0, 60, len = 4))+
  scale_fill_brewer(palette="Set2", direction = 1 ) +
  theme( 
         #title = element_text(face = "bold"),
         legend.title =element_blank(),
         legend.text = element_text(size = 5.75),
         axis.title.x= element_blank(),
         axis.title.y= element_text(family = "Old Standard TT, serif", size = 6.5),
         axis.ticks.x= element_blank(),
         axis.ticks.y = element_blank(),
        # panel.grid.major = element_blank(), 
       #  panel.grid.minor = element_blank(),
         legend.position = "top",
         plot.background = element_rect(fill = 'white'),
         panel.background = element_rect(fill = 'white')
  ) 
)%>%
  config(displayModeBar = FALSE)

```




### Transport emissions have grown more than any other sector, increasing nearly 60% since 1990.

```{r message=FALSE, warning=FALSE}


# Read data into per_capita_emission object
Emission_from_transport <- read.csv('Emission_from_transport.csv') 

Emission_from_transport$Transport <- reorder(Emission_from_transport$Transport, -Emission_from_transport$Average_emissions_per_kilometre.person_.gCO2..km.)

Emission_from_transport$Emissions_gCO2_per_person_per_km<-Emission_from_transport$Average_emissions_per_kilometre.person_.gCO2..km.

Emission_from_transport$Transport<-revalue(Emission_from_transport$Transport, c("Metro_train "="Metro Train", "Light_rail "="Light Rail"))
# Bar Plot
ggplotly(ggplot(data=Emission_from_transport, aes(x=Transport, y=Emissions_gCO2_per_person_per_km)) +
           geom_bar(stat="identity",fill='#8AA4E5') +
           #geom_text(aes(label = c("12g","13g","18g","184g")),
           #size = 5, hjust = 0.5, vjust = 0)+
           #  labs(title= "Emission from different forms of transport per km/person")+
           #geom_text(aes(label=Average_emissions_per_kilometre.person_.gCO2..km.),vjust=0, size=10)+
           theme(axis.text.x = element_text()) +
           xlab("Transport Mode")+ 
           ylab("gCO2/pkm")+
           theme( legend.title =element_blank(),
                  legend.text = element_blank(),
                  legend.position = "none",
                  axis.title.x = element_blank(),
                  axis.title.y= element_text(family = "Old Standard TT, serif", size = 6.5),
                  axis.ticks.x= element_blank(),
                  axis.ticks.y = element_blank(),
                  plot.background = element_rect(fill = 'white'),
                  panel.background = element_rect(fill = 'white'),
                  #panel.grid.major = element_blank(), 
                  panel.grid.minor = element_blank()
                  
                  
           ))%>%
  config(displayModeBar = FALSE)

```

Data Sources
============================================
[1] Greenhouse Gas Emissions - Sources of Greenhouse Gas Emissions - (2020) - https://www.epa.gov/ghgemissions/sources-greenhouse-gas-emissions [Accessed 14 June 2020]

[2] CO2 and Greenhouse Gas Emissions - Per capita CO2 emissions - (2019) -  https://ourworldindata.org/co2-and-other-greenhouse-gas-emissions [Accessed 14 June 2020]

[3] How Australia accounts for agricultural greenhouse gas emissions - (2020) - Agriculture is an important source of greenhouse gas emissions - https://www.agric.wa.gov.au/climate-change/how-australia-accounts-agricultural-greenhouse-gas-emissions [Accessed 14 June 2020]

[4] Fuel mix contributing to total primary energy supply - Australia 2013 and OECD average 2014 - (2016)  - https://soe.environment.gov.au/file/44236 [Accessed 14 June 2020]

[5] CO2 and Greenhouse Gas Emissions - Total greenhouse gas emissions - (2019) -  https://ourworldindata.org/co2-and-other-greenhouse-gas-emissions [Accessed 14 June 2020]

[6] Factsheet Transport Emission: Driving Down Car Pollution in Cities - Table 2: Emissions for different forms of transport - (2016) - https://www.climatecouncil.org.au/resources/transport-fact-sheet/https://www.climatecouncil.org.au/wp-content/uploads/2017/09/FactSheet-Transport.pdf [Accessed 14 June 2020]